### 2015-12-09 paperReviews.tex
getting a start

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Paper summaries of related work}
\author{}
\date{December 2015}

\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\maketitle

\section{Introduction}
Give a short description of papers that are useful.

\section{\cite{yin2011}}
Surveys a data set of approximately 500 (mis)configuration file and classifies the types of errors. 
Gives a frequency for each classification.
The frequency for the errors we are interested in are as follows: (a) = 15\%,...

\includegraphics[width=\textwidth]{configErrors.png}

\bibliographystyle{plain}
\bibliography{references}
\end{document}