### 2015-12-12 proposedSolution.tex
writing pseduocode

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Proposed Solution}
\author{}
\date{December 2015}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}

\definecolor{identifierColor}{rgb}{0.65,0.16,0.16}
\definecolor{comment_color}{rgb}{0.40,0.46,0.3}
\definecolor{num_color}{gray}{0.55}
\lstset{
  basicstyle=\footnotesize,
  breaklines=true,
  frame=bottomline,
  language=haskell,
  %identifierstyle=\color{identifierColor},
  morecomment=[l][\color{comment_color}\ttfamily]{--},
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  %captionpos=b,                    % sets the caption-position to bottom
  %commentstyle=\color{mygreen},    % comment style
  %frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  otherkeywords={*,let, Server, Replication, FaultGraph, rankRCG, print, fialProb, goal, ...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{num_color}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  %title=\lstname                  % show the filename of files included with \lstinputlisting; also try caption instead of title
  mathescape=true,
  tabsize=3,
  literate=*{->}{{\textcolor{blue}{$\to$}}}{1}
           {<-}{{\textcolor{blue}{$\leftarrow$}}}{1}
}

\begin{document}

\maketitle

\section{Introduction}
In order to solve the misconfiguration problems outlined in \cite{yin2011}, we take a two stage process.

\begin{enumerate}
  \item Preprocessor
  \begin{enumerate}
    \item Given a set of config files in a language $[C_{httpd}]$, convert them to a set $[D]$ in our language.
    \item Learn constraints on syntax based on $[D]$ to generate the rule set $R_{httpd}$
  \end{enumerate}
  \item Usertime checking
  \begin{enumerate}
    \item convert a given configuration file $c_{httpd}$, into a file $d$, in our own language.
    \item check $d$ against $R_{httpd}$ and report errors to user.
  \end{enumerate}
\end{enumerate}

Expressed in pseduocode:

\begin{lstlisting}
main = 
  r = preproc learningSet
  e = usertime r userFile

preproc :: [ConfigFile a] -> (RuleSet,a)
preproc cs =
  ds = map convert cs 
  rules = learnOn ds
  fileType = getType cs
  return (rules,fileType)

convert :: ConfigFile a -> ConfigFile Common
getType :: [ConfigFile a] -> a
learnOn :: [ConfigFile Common] -> RuleSet

usertime :: (RuleSet,a) -> ConfigFile a -> Maybe Error
usertime r c =
  d = convert c
  e = verifyOn r d
  return e

data RuleSet = {Lexical = [Clause], Syntax = [Clause], Value = [Clause]}
\end{lstlisting}

\section{Preprocessor}
The first step is to use a \texttt{CONVERTER} to translate any config file type (httpd,MySQL,etc.) to a common format. 
Our \texttt{LEARNER} for learning constaints will depend on all files having a common format. 
The types of constraints, and the proposed solution to learn such constraints are outlined here.

\begin{lstlisting}
learnOn :: [ConfigFile Common] -> RuleSet
learnOn cs = 
  rs = map genConstraints cs
  foldl1 mergeRules rs

genConstraints :: ConfigFile Common -> RuleSet
genConstraints c =
  r1 = learnLexicalConstraints c
  r2 = learnSyntaxConstraints c
  r3 = learnValueConstraints c
  return RuleSet {r1, r2, r3}

mergeRules :: RuleSet -> RuleSet -> RuleSet
mergeRules rs rs' =
  return RuleSet
    { mergeLex (Lexical rs) (Lexical rs')
    , mergeSyn (Syntax rs)  (Syntax rs')
    , mergeVal (Value rs)   (Value rs')}
\end{lstlisting}

\subsection{Format - Lexical}
This problem is exemplified in box (a) of the image - a certain parameter cannot contain lowercase values.
To build these rules, we collect every instance of setting that parameter over $[D]$, then infer a partition (like strisynth) on those examples.

\begin{lstlisting}
learnLexicalConstraints :: ConfigFile Common -> [Clause]
learnLexicalConstraints =

mergeLex =
  strisynth?
\end{lstlisting}

\subsection{Format - Syntax}
This problem is exemplified in box (b) and (c) of the image.
These are ordering/dependency errors and require analysis over multiple lines.

\begin{lstlisting}
learnSyntaxConstraints :: ConfigFile Common -> [Clause]
learnSyntaxConstraints c =
  ls = lines c
  concatMap (makeOrderPairs ls) ls

makeOrderPairs :: [Text] -> Text -> [(Text,Text)]
makeOrderPairs ls l = 
  map (,l) ls

mergeSyn :: [Clause] -> [Clause] -> [Clause]
mergeSyn curr new =
  combined = intersect curr new
  return combined
\end{lstlisting}

\subsection{Value - Value Inconsistency}
This problem is exemplified in box (f) (g) and (h) of the image.
I don't understand (f), how can we know if query.log is a file or a table.

Interger constraints are hard to find since we dont know which ones are supposed to be coorelated.
Maybe we can build constraints between all integers in a file, then remove constraints as we see counter examples.
Is this counter-example guided synthesis?

\begin{lstlisting}
learnValueConstraints :: Config Common -> RuleSet
learnValueConstraints c =
  ls = lines cs
  tys = seperateValsOnType ls :: [(Type,Text)]
  concatMap (makeRelations tys) tys

makeRelations :: [(Type,Text)] -> (Type,Text)
makeRelations ts (ty,tx) =
  map (f (ty,tx)) ts
  f (ty,tx) (ty',tx') = if ty==ty' then synthesizeComparison ty tx tx' else []

synthesizeComparison :: Type -> Text -> Text -> Clause
synthesizeComparison Int t t' =
  if t<t' then "t<t'"
  if t==t' then "t==t'"
  ...
synthesizeComparison String t t' =
synthesizeComparison IP t t' =
...

mergeVal rs rs' = 
\end{lstlisting}

\section{Usertime Checking}

\bibliographystyle{plain}
\bibliography{references}
\end{document}