### 2016-01-25 intro.tex

\section{Introduction}

Configuration errors are one of the most important root-causes of 
modern software system failures~\cite{xu15systems,yin11anempirical}. 
In practice, misconfiguration problems may result in 
security vulnerabilities,
application crashes, severe disruptions in software functionality,
and incorrect program executions%
~\cite{xu15systems,zhang14encore,yuan11context}.
Although several efforts have been proposed to automate configuration
error diagnosis after failures occur%
~\cite{wang04automatic,attariyan10automating,%
su07autobash,whitaker04configuration}, 
they rely on manual ways to understand and detect the failure symptoms.
Such solutions typically incur high administration cost (\eg, 
requiring users to write pattern-matching scripts) 
and are unreliable (\eg, some errors may show no user-visible symptoms).
These shortcomings often lead to prolonged delays between
the occurrence and the detection of errors, thus causing unrecoverable
damage to system states.

We expect to explore a fundamentally different means to tackle 
the misconfiguration problem. Rather than using retroactive approaches
to deal with configuration errors after the software failures occur, 
we would like to {\em proactively} prevent misconfiguration problems 
{\em before} users first deploy the configuration files to 
their system settings. In other words, we aim to offer users 
a configuration error filter in order to significantly minimize 
the probability of misconfigurations in the system runtime.

Based on the above intuition, this paper presents \app, 
a proactive misconfiguration checker which takes users' 
configuration files as input and then reports potential configuration 
errors to the users before they deploy their systems 
according to these configurations.
At the heart of \app lies a machine learning mechanism, 
which learns rules and constraints from pluggable 
correct sampling configuration files (\ie, for the training purposes), 
and then detects errors violating the learned constraints.

\app mainly offers users three significant benefits.
First, there is no any extra burden added to users who only need to 
simply use \app as a ``seamless'' pre-checker for their configurations.
Second, because many misconfiguration errors have been eliminated 
by \app, the workloads of post-failure forensics in runtime
are significantly reduced, thus making these tools truly practical.
Finally, since there have been many correct but not specific 
example configuration files in practice, 
increasingly more samples can be freely added to \app,
thus giving \app evolutionary capability.

Using these ideas, we make the following contriubtions:

\begin{enumerate}

  \item Our tool, \app can learn a language model from a set of examples of correct configuration files, and use the model to verify new files.
  \item We present a probabalistic type that can be used to assign a confidence distribution over a set of types to a valule.
  \item In \app, we define a minimal interface for describing a verification attribute in a learning context, making it easy to add new rules to the system.

\end{enumerate}