### 2016-01-28 design.tex
a pretty good outline for the meat of the paper

\section{\app Design}

\xxx{
\app Design: 
  \begin{itemize}
  \item Present an architecture of \app with a fig. Briefly describe 
    how it works (step by step).
  \item Detail learner part
  \item Merge
  \item Check 
  \item Limitations
  \end{itemize}
}

\section{Learning strategy}

A primary concern in any machine learning type task is to minimize false negatives and false positives.
In the context of configuration file verification,
  a false positive is when \app reports an error on a valid confiuration file and
  a false negative is when \app fails to reports an error on an invalid confiuration file.
Too many false positives will cause users to ignore the reported error\cite{}.
However, since the cost of system failure is so high from a misconfiguration, \app propritzes the minimization of false negatives.

While a traditional classififcation learning machine learning approach can reduce both of these situations, there is can generally be no garuntee that all false negatives will be eliminated.
Instead of building classification models over the learning set (such as an SVM), we learn the largest set of rules that all correct configuration files satisfy.
In this way, \app can garuntee that, over the set of rules we consider, there will be no false negatives that could have been caught with the given learning set.
The only case of a false negative can be when there was no evidece of such a rule in the learning set - we cannot generate rules from nothing.
Framed as the question "Is this file valid", \app is complete but not sound \xxx{maybe? Im not sure}

That is, taking the following definitions:
\begin{multline*}\\
\text{\{Considered Rules\}} = \forall files \in \text{\{Learning Set\}}, \{ r | r(files) = True \land r \text{ is non-trivial}\}\\
\text{\{Reported Rules\}} = {r | r(userfile)=False } \\
\text{\{True Rules\}} = \{r | \text{ if } r(userfile)=False \text{ then system crash } \land \\
   \exists file \in \text{\{Learning Set\}}, r(file) \text{ is non-trivial}\}
\end{multline*}

We have the following specification of \app.
\begin{multline*}\\
\text{\{True Rules\}} \subseteq \text{\{Considered Rules\}} \\
\forall r \in \text{\{True Rules\}}, if r(userfile)=False then \exists r \in \text{\{Reported Rules\}} \\
\neg \forall r \in \text{\{Reported Rules\}}, r \in \text{\{True Rules\}} \\
\end{multline*}

Another benefit to a rule based approach is that, unlike many classificaton models, \app can actually report the reason for failure, similar to a comiler for a programming language.
This is in contrast to neural nets for example, where we would just get a boolean value, and the mechanics of the system are entirly lost.
Reporting useful errors that specify a point of failure is important to help users fix their misconfigurations.

\section{Intermediate Representation}

Since we don't want to rewrite checkers for every config file type, we made an intermediate representation.
The intermediate reprentation is simply a list of keyword value pairs.
A keyword is the setting being adjust in the configuration file, and the value is the new value for that setting.
The user only needs to specify the delimiters of the file (characters for assignment and comments usually).

\section{Rules}

In order to flexibly consider many different types of rules, we introduce a inferface (or typeclass in Haskell) that each rule type must support.
The typeclass is presented here with simplified type signatures for those unfamiliar with Haskell.

\begin{lstlisting}
class Attribute r where
  learn :: IRConfigFile -> Set r
  merge :: Set r -> Set r -> Set r
  check :: Set r -> IRConfigFile -> Error
\end{lstlisting}

To build the set Considered Rules, we first collect all possible rules for every file.
Then we merge the all the rules to create our final set.

\subsection{learn}
  For a single given file in the intermediate representation format, learn a set of rules.
  In fact, we need to learn every possible rule that this file can satisfy.
  By overfitting to each file, we can eventually garuntee the completness of \app

\subsection{merge}
  Merging the sets of rules from two files to build a new set that is true over both files
  This is generally an intersection like operation, but depending on the particular rule, there may be different stratgeies for merging.
  The formal specification of this method is that:
  \begin{multline*}
  \text{Output Set} = \{r \mid \\
    r \in \text{\{Input Set1\}} \cup \text{\{Input Set2\}} \land \\
    \exists file \forall r' \in \text{\{Input Set1\}} \cup \text{\{Input Set2\}}, r(file) = True \land r'(file) = True \} \\
  \end{multline*}

\subsection{check}
  To check a file by using a rule set, we simply take all the rules that are releveant to the user's file.
  Rules that are relavent are the ones where both parts of the ordering are present.
  We learn the rule set for the user file, and every rule in the learned set must be present in the user file.

\section{Implementation}

\subsection{System}
Since we learn a set of rules on each file in isolation from the other, we have an embarresingly parallel situation.
Haskell allows us to easily take advantage of by using the parallel mapping library, parmap.

\begin{lstlisting}
  potentialRules = parmap findAllRules learningSet.
  finalRules = foldl1 mergeRules potenialRules
\end{lstlisting}

\subsection{Type Error Rules}
This builds a map between keywords and types, using the values as evidence.
see quantum.tex for more
i will write more on this tomorrow.

\subsection{Integer Relation Rules}
we only consider the relations , (==), (<=), (>=) because they are easy to pass aroud (as actual functions) in haskell.
We also committed cardinal since and create an instance for equality over these functions.
instance Eq (Int->Int->Bool) where

With more engineering effort, this could be extened with the use of a SMT solver to create more fine grained relational rules.
From our experience, more specific rule are not really needed for integer relations in configuration files.
However an SMT solver approach would be particularly useful for relations on strings, especially when considering substring relations between filepaths.

\subsection{Ordering and Missing Entry Rules}
These are the simpiliest of all rules, just making a lot of pairs and seeing which pairs continue to appear over the learning set.