### 2016-01-29 system.tex
lala

\section{System Overview}
\label{sec:system}

\begin{figure}[t] \centering
\includegraphics[width=0.8\textwidth]{figs/overview}
\caption{\app's workflow overview.}
\label{fig-overview}
\end{figure}

Fig.~\ref{fig-overview} describes an overview of our system. Our goal is first to learn

Initially
we use a large number of correct configuration files as training data