### 2016-01-29 evaluation.tex
added eval

\section{evaluation}

We report error that was unrelated to the value of interest as a false positive.
It is worth noting that this is in fact a conservative estimate.
Since these benchmarks are taken from online forums, there is no garuntee the files contain only a single error.
Indeed, while running these benchmarks, \app found errors in the file that were similar to other benchmarks.

We fail one benchmark in Value Relations because we do not yet support relations between file sizes of different units (Mb to Kb)

\begin{table}[]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{lll}
Error Type       & Passing Tests & False Positives  \\
Missing Entry    & 5/5           & 1,0,0,0,1        \\
Type Error       & 4/4           & 0,0,0,0          \\
Keyword Ordering & 4/4*          & 0,10,1,0         \\
Value Relations  & 4/5           & 0,0,0,0,0        \\
*One passing test gives a type error on the misconfigured value rather than an ordering error, as it should.
\end{tabular}
\end{table}