### 2016-01-29 system.tex

\section{Learning the Rules}
\label{sec:system}

Before describing how our system for learning rules works, we first outline a translator which converts configuration files into an intermediary representation. Typing is based on an
already given set of basic types. However, typing is also a system module than can be easily extended to support
more types. In that case the user will need to provide rules for type inference and probability distributions for values where type inference is ambiguous.

\subsection{From Untyped Expressions to Quantum Types}

Configuration files usually consist of variable assignments, or some simple instructions about including files or libraries. We initially
collect all assignments into a set of pairs $(n, v)$, where $n$ denotes 
variable name and $v$ denotes variable value. However, based only on the 
value $v$ we cannot determine the type of $n$.

Consider for instance the following example:\\
\texttt{\hspace*{2em}foo = 300\\
\hspace*{2em}bar = 300.txt}\\
Most likely \texttt{foo} is an integer, but it could also be a string.
In the latter case we want to learn the rule $ \texttt{foo} \in \textsf{substrings}(\texttt{bar})$. We introduce {\emph{quantum types}} to 
address this issue.

Let $\mathcal{T}$ be a set of basic types. In \app set $\mathcal{T}$ contains strings, integers, file paths, sizes and IP addresses. 
A quantum type built from $\mathcal{T}$ is a list of pairs $[(\tau_1, p_1),\ldots,(\tau_n, p_n)]$ such that $\tau_i \in \mathcal{T}$, 
$0 \le p_i \le 1$ 
and $\Sigma p_i = 1$. These probabilities are learned from the training 
corpus.

When a value has a quantum type, we generate rules for all its types. This means that by assigning {\texttt{foo}} a quantum type 
(e.g. $(\texttt{foo}, 300, [(\textsl{Int},90\%),(\textsl{String},10\%)])$
we now generate rules for strings and integers.

For a given value, its quantum type is updated over time. When the type inference uniquely determines the type, the probability of all other types is set to zero, and  the associated rules are withdrawn.

This idea is closely related to existentially quantified types \cite{Launchbury93lazyfunctional}.