### 2016-01-29 related.tex

\section{Conclusions}

In this paper, we propose a highly-modular framework that helps 
users to verify their configuration files through learning
rules from datasets of correct configuration files.
We have implemented design as a tool, named \app, in Haskell 
and evaluated \app by checking real incorrect configuration file datasets.