### 2016-01-29 system.tex
reduce type section

\section{Learning the Rules}
\label{sec:system}

Before describing how our system for learning rules works, we first outline a translator which converts configuration files into an intermediary representation. Typing is based on an
already given set of basic types. However, typing is also a system module than can be easily extended to support
more types. In that case the user will need to provide rules for type inference and probability distributions for values where type inference is ambiguous.

\subsection{From Untyped Expressions to Probabilistic Types}

Each line of a configuration files can be reduced to a keyword value pair
In the parsing step we collect all lines in the file into a set of such pairs $(k, v)$, where $k$ denotes 
variable name and $v$ denotes variable value. 
However, based only on a single example value of $v$ we cannot fully determine the type of $k$.
We instead assign a distribution of type to a value, an idea is closely related to existentially quantified types \cite{Launchbury93lazyfunctional}.

Consider for instance the following example:\\
\texttt{\hspace*{2em}foo = 300\\
\hspace*{2em}bar = 300.txt}\\
Most likely \texttt{foo} is an integer and we learn an equality rule, but it could also be a string.
In this case we want to learn the rule $ \texttt{foo} \in \textsf{substrings}(\texttt{bar})$. 
For more advanced rules, we may need to even compose these types to learn complex rules.
We introduce {\emph{probabilistic types}} to address this issue.

Let $\mathcal{T}$ be a set of basic types. In \app set $\mathcal{T}$ contains strings, integers, file paths, sizes and IP addresses. 
A probabilistic type built from $\mathcal{T}$ is a list of pairs $[(\tau_1, p_1),\ldots,(\tau_n, p_n)]$ such that $\tau_i \in \mathcal{T}$, 
$0 \le p_i \le 1$ 
and $\Sigma p_i = 1$. These probabilities are updated each time a new example value for a keyword is encountered.
When a value has a probabilistic type, we generate rules for all its types. This means that by assigning {\texttt{foo}} a probabilistic type 
(e.g. $(\texttt{foo}, 300, [(\textsl{Int},90\%),(\textsl{String},10\%)])$
we now generate rules for both strings and integers.
Once the type inference can uniquely determine the type, the probability of all other types is set to zero, and the associated rules are withdrawn.