### 2016-01-29 related.tex
autocommit

\section{Conclusions}

In this paper, we propose a highly-modular framework that allows verification of configuration files, even without language model of the file.
\app can easily be extended to support new verification properties and new configuration langauages.
New verification properties require only a small amount of code and are not language specific, all indicating that \app could be widely adopted by system administrators.
Having such verification tool that scales in both performance and expressivity can revolutionize configuration file checking, reducing the cost of system maintaince and failure dramatically.