### 2016-01-29 related.tex
autocommit

\section{Conclusions}

In this paper, we introduce \app, a highly-modular framework that allows verification of configuration files, even without a language model of the file.
New verification properties require only a small amount of code and are not language specific, all indicating that \app could be widely adopted by system administrators.
Such a verification tool that scales in both performance and expressivity can revolutionize configuration file checking, reducing the cost of system maintaince and failure dramatically.