### 2016-05-01 OSDI/check.tex
config: for osdi16

\section{Checker}

\ennan{Here, we need to describe how we check whether 
the target configuration file meets learned rules using SMT solver.}