### 2016-05-01 OSDI/limit.tex

\section{Discussion and Limitations}

\ennan{Of course, our tool has many limitations. For example, we cannot
detect the misconfiguration correlation across multiple individual
systems unwilling to share configuration information with each other.}