### 2016-05-01 OSDI/limit.tex
config: related work and limitation sec

\section{Discussion and Limitations}

\ennan{Of course, our tool has many limitations. For example, we cannot
detect the misconfiguration correlation across multiple individual
systems unwilling to share configuration information with each other.}

This section discusses a few \app's limitations
and possible solutions.

\para{Legal misconfigurations.}
While \app can check diverse configuration errors without
human participations, all the misconfiguration detection techniques,
including \app, cannot handle some types of configuration errors
resulting from events occurred during system runtime.
Such configuration errors are referred to as {\em legal misconfigurations}%
~\cite{yin11anempirical}. As shown in Yin {\em et al.}%
~\cite{yin11anempirical}, many parameter misconfigurations have 
perfectly legal parameters but do not deliver the functionality intended
by users. These cases are more difficult to detect by
automatic checkers and may require more user training or
better configuration design.
A potential solution is to combine existing misconfiguration diagnosis
tools, \eg, X-ray~\cite{}, with \app in order to enhance the 
misconfiguration checking capability.

\para{}