### 2016-05-07 OSDI/learner-ranges.tex
add intuition for ranges

\section{Ranges}
\label{sec-learn-ranges}

We discuss the technique we use to detect anomalous values for numerical attributes. For a given 
test file, let $A$ be the set of attributes contained in the file. Let $A_n$ be the subset of attributes 
of $A$ which are numerical types. Then, for each attribute $a \in A_n$, we construct a vector $v_a$ of the 
values corresponding to attribute $a$ seen over the entire data-set. For each $v_a$, we compute 
an interval  $$[\hat{v_a} - 3*MAD(v_a), \hat{v_a} + 3*MAD(v_a)],$$ 
where $\hat{v_a}$ represents the median over the values in $v_a$ and $MAD(v_a$) refers to the 
median absolute deviation. This is a standard outlier detection test, namely the Hampel identifier. 
\footnote{Mathematically, $MAD(v_a) = 1.4826* median(|v_a - \hat{v_a}|)$, estimating standard deviation 
for a normal distribution.} If the value for a numerical attribute in a test file 
falls outside of this interval, a warning is printed to the user indicating the violating value, 
the attribute, and the upper or lower Hampel threshold.

The intuition for this test is that if the user has input a value that falls outside of an 
interval containing values that are considered "normal" over the entire data set, 
that value will probably cause an error, in particular for performance. We cannot know for sure
if this value will cause an issue. For instance, a user might have a machine with 
particularly high-end hardware, in which case a value beyond the upper Hampel threshold may be appropriate.