### 2016-05-09 OSDI/eval.tex
config: eval and learner

\section{Implementation}

\ennan{We need to describe how do we implement each module. We may need to
have three paragraphs: one is for translator, one is for learner 
and one is for checker.}

\begin{table*}[t]
\centering
\caption{Sampled benchmarks for misconfiguration detection}
\label{table-casestudy}
\begin{small}
\begin{tabular}{|l|l|l|l|}
\hline
{\bf ID} & {\bf Problem Description} & {\bf Error Type} & 
{\bf \app Report}  \\ 
\hline
\hline
1 & MySQL cannot create test file  
& Entry Missing 
& Expected innodb\_force\_recovery=1 in the \\ & & 
&  same file as: innodb\_strict\_mode=1\\ \hline

2 & How can my wait\_timeout/interactive\_timeout  
& Entry Missing
& Expected set\_time=1 in the same\\ & be ignored 
& & file as wait\_timeout \\ \hline

3 & MySQL on EC2 instance becomes very slow
& Correlation
& Expected key\_buffer\_size $>=$ \\ & after the first query
&& thread\_size * sort\_buffer\_size \\ \hline

4 & MySQL cannot successfully make  
& Type Error
& Expected master-host[mysqld] should 
 \\ & replication &&  be a valid IP address\\ \hline

5 & MySQL running on production servers has
& Type Error
& Expected set-variable[myisamchk] \\ &  gone away error in memory
&&  should be a memory size like 20M \\ \hline

6 & Cannot start MySQL 5.5 as normal 
& Ordering
& Expected socket[mysqld]=/var/lib/ \\ & user Fedora 15 
&&  mysql/mysql.sock BEFORE  \\ & && user[mysqld]=mysql

\\ \hline

7 & Fail to login MySQL
& Type Error
& Expected old\_passwords[mysqld] 
 \\ &&&  should be a valid value \\ \hline

8 & MySQL running at CentOS cannot know 
& Ordering
& Expected localhost before bind-
 \\ & enable named pipe
&&  address[mysqld] \\ \hline

9 & Troubles installing MySQL5 via  Darwin 
& Ordering
& Expected port[mysqld]=3306 BEFORE \\ & Ports
&&  socket[mysqld]=/tmp/mysql.sock \\ \hline

10 & Fail to install Percona xtraDB 
& Entry Missing
& Expected wsrep\_node\_address[mysqld] \\ & cluster on ubuntu 13.04
&&  in the same file as: wsrep\_sst\_method \\ &&& [mysqld] \\ \hline

11 & MySQL access denied for user rootlocalhost 
& Ordering
& Expected port[mysqld]=3306  \\ & &&   BEFORE socket[mysqld]= \\ &&& /tmp/mysql.sock
\\ \hline

12 & MySQL Partition Problem
& Correlation
& Expected innodb\_flush\_max  \\ & &&  \_commit[mysqld] $>=$ innodb \\ &&&  \_support[mysqld] \\ \hline

13 & mysql\_upgrade script problems on 
& Type Error
& Expected log-slow-queries[mysqld]  \\ & MySQL 5.0.24 
&&  should be a path type\\ \hline

14 & MySQL: bug report!!
& Ordering
& Expected innodb\_file\_io\_threads[mysqld]  \\ & 
&& BEFORE innodb\_log\_files\_in\_group[mysqld]\\ \hline

15 & MySQL max connections changed 
& Correlation
& Expected innodb\_file\_per\_table \\ & without notice
&&  [mysqld] $<=$ max\_connections[mysqld]  \\ \hline

16 & Enabling log in MySQL 5 6 
& Type Error
& Expected sql\_mode[mysqld] should  \\ & prevents server from starting
&&  be a string \\ \hline

17 & MySQL needs help for rapidly growing 
& Correlation
& Expected innodb\_buffer\_pool\_size \\ & table and decreasing speed
&&  [mysqld] $>$ sort\_buffer\_size[mysqld] \\ \hline

18 & MySQL filled all RAM on the system
& Correlation
& Expected join\_buffer\_szie * \\ & && max\_connections $<=$ key\_buffer\_size \\ \hline

19 & MySQL hit a max limit
& Correlation
& Expected join\_buffer\_size * \\ &&& max\_connections $<=$ key\_buffer\_size \\ \hline

20 & Cannot create a new thread
& Correlation
& Expected join\_buffer\_size * \\ &&& max\_connections $<=$ key\_buffer\_size \\ \hline

\end{tabular}
\end{small}
\end{table*}

\section{Evaluations}

We mainly conduct three parts of experiments to evaluate our \app
prototype. We want to answer several questions:

\begin{itemize}

\item Whether \app can detect real-world configuration errors?

\item How does the size of training dataset impact on the correctness
  of \app?

\item What is the rule inference accuracy?

\item How long can we run the verification?

\end{itemize}

\subsection{\app's effectiveness}

In order to evaluate the effectiveness of \app, we extract
20 misconfiguration files from a MySQL dataset~\cite{xu15hey},
and perform \app on them. \ennan{Here, we first need a table to 
list each configuration error with error type, description,
and whether \app correctly reports the issues.}

Next, we inject some errors to correct configuration files,
and run \app to check whether our framework can correctly detect
these injected errors. \ennan{We also need a table here, 
and list each injected problem.}

Finally, we check suspicious values.

\begin{table*}[t]
\centering
\caption{Sampled benchmarks for anomaly detection}
\label{table-anomaly}
\begin{small}
\begin{tabular}{|l|l|l|l|}
\hline
{\bf ID} & {\bf Problem Description} & {\bf URL} & 
{\bf \app Report}  \\ 
\hline
\hline
1 & MySQL Server has gone away error in Wamp 
& goo.gl/axnezi  
& WARNING: Violated Upper Hampel Rule for 
\\ &  & 
& {\tt max\_allowed\_packet} with value 104857600 
 \\ \hline

2 &  MySQL has abnormally high load for CPU 
& goo.gl/JrRLrR
& WARNING: Violated Upper Hampel Rule for  
\\ & & 
& {\tt sort\_buffer\_size} with value 1048576000 \\
& & & WARNING: Violated Upper Hampel Rule for  
\\ & & 
& {\tt read\_rnd\_buffer\_size} with value 283115520 \\ \hline

3 & User is having performance issues with MySQL 
& goo.gl/34jTB5
& WARNING: Violated Upper Hampel Rule for  
\\ & & 
& {\tt query\_cache\_limit} with value 134217728 \\ \hline

\end{tabular}
\end{small}
\end{table*}

\subsection{Training Set Impact}

In this experiment, we vary the number of training configuration files,
and observe whether the correctness of \app would increase 
accordingly. \ennan{We may need a table whose x-axis is the number
of entries in the training configuration files, and y-axis is the 
accuracy or something close.}

\subsection{The Run-Time of Verification}

\ennan{Here, we need three pictures. In the first figure, x-axis should
be the number (or the size) of entries in the training dataset, and 
y-axis should be the run-time of parsing. In the second figure,
we need to measure the time of generating rules. }