### 2016-05-09 OSDI/table.tex

\begin{table*}[tbp]
\centering
\caption{Sampled benchmarks for misconfiguration detection}
\label{table-casestudy}
\begin{small}
\begin{tabular}{|l|l|l|l|}
\hline
{\bf ID} & {\bf Problem Description} & {\bf Error Type} & 
{\bf \app Report}  \\ 
\hline
\hline
1 & MySQL cannot create test file  
& Entry Missing 
& Expected innodb\_force\_recovery=1 in the \\ & & 
&  same file as: innodb\_strict\_mode=1\\ \hline

2 & How can my wait\_timeout/interactive\_timeout  
& Entry Missing
& Expected set\_time=1 in the same\\ & be ignored 
& & file as wait\_timeout \\ \hline

3 & MySQL on EC2 instance becomes very slow
& Correlation
& Expected key\_buffer\_size $>=$ \\ & after the first query
&& thread\_size * sort\_buffer\_size \\ \hline

4 & MySQL cannot successfully make  
& Type Error
& Expected master-host[mysqld] should 
 \\ & replication &&  be a valid IP address\\ \hline

5 & MySQL running on production servers has
& Type Error
& Expected set-variable[myisamchk] \\ &  gone away error in memory
&&  should be a memory size like 20M \\ \hline

6 & Cannot start MySQL 5.5 as normal 
& Ordering
& Expected socket[mysqld]=/var/lib/ \\ & user Fedora 15 
&&  mysql/mysql.sock BEFORE  \\ & && user[mysqld]=mysql

\\ \hline

7 & Fail to login MySQL
& Type Error
& Expected old\_passwords[mysqld] 
 \\ &&&  should be a valid value \\ \hline

8 & MySQL running at CentOS cannot know 
& Ordering
& Expected localhost before bind-
 \\ & enable named pipe
&&  address[mysqld] \\ \hline

9 & Troubles installing MySQL5 via  Darwin 
& Ordering
& Expected port[mysqld]=3306 BEFORE \\ & Ports
&&  socket[mysqld]=/tmp/mysql.sock \\ \hline

10 & Fail to install Percona xtraDB 
& Entry Missing
& Expected wsrep\_node\_address[mysqld] \\ & cluster on ubuntu 13.04
&&  in the same file as: wsrep\_sst\_method \\ &&& [mysqld] \\ \hline

11 & MySQL access denied for user rootlocalhost 
& Ordering
& Expected port[mysqld]=3306  \\ & &&   BEFORE socket[mysqld]= \\ &&& /tmp/mysql.sock
\\ \hline

12 & MySQL Partition Problem
& Correlation
& Expected innodb\_flush\_max  \\ & &&  \_commit[mysqld] $>=$ innodb \\ &&&  \_support[mysqld] \\ \hline

13 & mysql\_upgrade script problems on 
& Type Error
& Expected log-slow-queries[mysqld]  \\ & MySQL 5.0.24 
&&  should be a path type\\ \hline

14 & MySQL: bug report!!
& Ordering
& Expected innodb\_file\_io\_threads[mysqld]  \\ & 
&& BEFORE innodb\_log\_files\_in\_group[mysqld]\\ \hline

15 & MySQL max connections changed 
& Correlation
& Expected innodb\_file\_per\_table \\ & without notice
&&  [mysqld] $<=$ max\_connections[mysqld]  \\ \hline

16 & Enabling log in MySQL 5 6 
& Type Error
& Expected sql\_mode[mysqld] should  \\ & prevents server from starting
&&  be a string \\ \hline

17 & MySQL needs help for rapidly growing 
& Correlation
& Expected innodb\_buffer\_pool\_size \\ & table and decreasing speed
&&  [mysqld] $>$ sort\_buffer\_size[mysqld] \\ \hline

18 & MySQL filled all RAM on the system
& Correlation
& Expected join\_buffer\_szie * \\ & && max\_connections $<=$ key\_buffer\_size \\ \hline

19 & MySQL hit a max limit
& Correlation
& Expected join\_buffer\_size * \\ &&& max\_connections $<=$ key\_buffer\_size \\ \hline

20 & Cannot create a new thread
& Correlation
& Expected join\_buffer\_size * \\ &&& max\_connections $<=$ key\_buffer\_size \\ \hline

\end{tabular}
\end{small}
\end{table*}