### 2016-05-10 OSDI/impl.tex
config: impl

\section{Implementation}

\ennan{We need to describe how do we implement each module. We may need to
have three paragraphs: one is for translator, one is for learner 
and one is for checker.}