### 2016-05-10 OSDI/eval.tex
config: first part in eval

\section{Evaluations}
\label{sec-eval}

We conduct three parts of experiments to evaluate our \app
prototype. We want to answer several questions:

\begin{itemize}

\item Whether \app can detect real-world configuration errors?

\item How does the size of training dataset impact on the correctness
  of \app?

\item What is the rule inference accuracy?

\item How long can we run the verification?

\end{itemize}

\input{table}

\subsection{\app's Effectiveness}

We now apply \app to check against real-world misconfiguration problems.
We use a real misconfiguration dataset~\cite{xu15hey}
that are collected from several online forums, \eg, 
Stack Overflow and MySQL forum.
Each configuration file in this dataset was posted by 
programmers or system administrators, 
when they confront the misconfiguration issues in practice.
This dataset contains 261 incorrect MySQL configuration files.

We apply \app to learn and check the configuration files in this
dataset. Then, we observe how many errors \app can detect,
and manually check whether these detected errors are indeed
configuration errors.
For all the 261 configuration files, we found \app is
able to report all the errors correctly.
For 217 of the 261 configuration files, we found \app's 
error detection results are correct and accurate,
\ie, correctly report the real misconfiguration problem.
However, for 44 of them, \app outputs relatively high false positives,
\eg, higher than 50 false positives,
which is not very helpful to users in practice.
The reason is our training set contains too many incorrect configuration
files, which produce a lot of noises to our learning process,
thus resulting in so many false positives.

In order to clearly evaluate the effectiveness of \app,
we extract 20 incorrect configuration files from the dataset.
As a case study, we present the problem descriptions of these files, 
the types of their errors, and the outputs produced by \app.
Table~\ref{table-casestudy} details this information.

As shown in Table~\ref{table-casestudy},
we found \app is capable of detecting many configuration errors previous
efforts, \eg, EnCore~\cite{zhang14encore}, cannot deal with.
To the best of our knowledge, no existing effort
is able to detect fine-grained value correlation errors in the 
MySQL configuration files.
For example, the third case in Table~\ref{table-casestudy}
presents \app detects the fine-grained correlation,
{\tt key\_buffer\_size} should be larger than 
{\tt sort\_buffer\_size} * {\tt max\_connections}.

\begin{table*}[t]
\centering
\caption{Sampled benchmarks for anomaly detection}
\label{table-anomaly}
\begin{small}
\begin{tabular}{|l|l|l|l|}
\hline
{\bf ID} & {\bf Problem Description} & {\bf URL} & 
{\bf \app Report}  \\ 
\hline
\hline
1 & MySQL Server has gone away error in Wamp 
& goo.gl/axnezi  
& WARNING: Violated Upper Hampel Rule for 
\\ &  & 
& {\tt max\_allowed\_packet} with value 104857600 
 \\ \hline

2 &  MySQL has abnormally high load for CPU 
& goo.gl/JrRLrR
& WARNING: Violated Upper Hampel Rule for  
\\ & & 
& {\tt sort\_buffer\_size} with value 1048576000 \\
& & & WARNING: Violated Upper Hampel Rule for  
\\ & & 
& {\tt read\_rnd\_buffer\_size} with value 283115520 \\ \hline

3 & User is having performance issues with MySQL 
& goo.gl/34jTB5
& WARNING: Violated Upper Hampel Rule for  
\\ & & 
& {\tt query\_cache\_limit} with value 134217728 \\ \hline

\end{tabular}
\end{small}
\end{table*}

\subsection{Training Set Impact}

In this experiment, we vary the number of training configuration files,
and observe whether the correctness of \app would increase 
accordingly. \ennan{We may need a table whose x-axis is the number
of entries in the training configuration files, and y-axis is the 
accuracy or something close.}

\subsection{The Run-Time of Verification}

\ennan{Here, we need three pictures. In the first figure, x-axis should
be the number (or the size) of entries in the training dataset, and 
y-axis should be the run-time of parsing. In the second figure,
we need to measure the time of generating rules. }