### 2016-05-10 OSDI/main.tex

\documentclass[letterpaper,twocolumn,10pt]{article}

\usepackage{times}            % standard fixed width font
\usepackage{usenix}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage{algorithmic}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{footnote}
\usepackage{cite}
\usepackage{amsfonts}
\usepackage{subfig}
%\usepackage{natbib}
\usepackage{hhline}
%\usepackage{multirow}
\usepackage{setspace} 
\usepackage{epsfig}
\usepackage[hyphens]{url}
\usepackage[colorlinks,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage[hyphenbreaks]{breakurl}
\usepackage{booktabs}
%\usepackage[compact]{titlesec}
%\usepackage{xcolor}
%\usepackage[algoruled,vlined,ruled,linesnumbered]{algorithm2e}
\usepackage{lipsum}
\usepackage{listings}

\lstset{
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  frame=bottomline,
  language=haskell,
  %identifierstyle=\color{identifierColor},
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  %captionpos=b,                    % sets the caption-position to bottom
  %commentstyle=\color{mygreen},    % comment style
  %frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  otherkeywords={*,let, Server, Replication, FaultGraph, rankRCG, print, fialProb, goal, ...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  %title=\lstname                  % show the filename of files included with \lstinputlisting; also try caption instead of title
  mathescape=true,
  tabsize=3,
  literate=*{->}{{\textcolor{blue}{$\to$}}}{1}
           {<-}{{\textcolor{blue}{$\leftarrow$}}}{1}
}

\usepackage[T1]{fontenc}
%Ruzica\usepackage[scaled=0.78]{DejaVuSansMono}

\clubpenalty=10000      % penalty for creating a club line at end of line.
\widowpenalty=10000     % penalty for creating a widow line at top of page.

% Select one or other if want to see comments.
% \com is sometimes displayed during draft.
\long\def\com#1{}
%\long\def\com#1{{\bf \sc comment: }{\small [#1]}{\bf \sc\ endcomment}\newline}

\long\def\ennan#1{{\color{blue}{\bf Ennan: }{\small [#1]}}}
\long\def\markk#1{{\color{violet}{\bf Mark: }{\small [#1]}}}
\long\def\ruzica#1{{\color{red}{\bf Ruzica: }{\small [#1]}}}
\long\def\jon#1{{\color{orange}{\bf Jon: }{\small [#1]}}}
%\long\def\xxx#1{}

% Use this macro to force page breaks where ugly widows/orphans occur;
% be sure to recheck all uses after any significant change to the text!
\def\widowpage{\pagebreak}

% Choose abbreviated or long-version alternatives in paper
%\long\def\abbr#1#2{#1}			% abbreviated version
\long\def\abbr#1#2{#2}			% long version

% Choose abbreviations or long names/titles in bibliography
%\def\bibbrev#1#2{#1}			% short version
%\def\bibbrev#1#2{#2}			% long version
\def\bibbrev#1#2{\abbr{#1}{#2}}		% follow abbr macro

% Abbreviated or full citation lists: \abcite{basic}{others}
\newcommand{\abcite}[2]{\abbr{\cite{#1}}{\cite{#1,#2}}}

% Conference abbreviations: \bibconf[Nth]{SOSP}{Symposium on ...}
\newcommand{\bibconf}[3][]{#1 \bibbrev{#2}{#3 (#2)}}

\newcommand{\ie}{{\em i.e.\xspace}}
\newcommand{\eg}{{\em e.g.\xspace}}

% system related terms
\newcommand{\app}{ConfigV\xspace}

% Fault graph terms

\newcommand{\para}[1]{\smallskip\noindent {\bf #1}}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{An Automatic Verification Framework for Software Configurations}

\author{~}

\maketitle
%\vspace{-50pt}

\input{abs}

\input{intro}
\input{motiv}
%\input{system}
\input{arch}
\input{trans}
\input{learner}
%\input{learner-ranges}
\input{check}
\input{limit}
\input{impl}
\input{eval}
\input{rel}
\input{concl}

%\newpage

\bibliographystyle{plain}
\bibliography{os}

\end{document}