### 2016-05-10 OSDI/concl.tex
config: concl

\section{Conclusion}

In this paper, we introduce \app, a highly modular framework 
that allows automatic verification of configuration files.
\app employs a translator to parse a sample configuration dataset
into a well-structured and typed intermediate representation,
and then uses a learner module to derive rules, 
thus building a language model.
For a given configuration file we want to verify,
\app uses the generated language model to check
whether any rule is violated.
We evaluate \app using a real-world dataset~\cite{configdataset}
which contains 261 incorrect MySQL configuration files.
Our experimental result shows \app is able to
correctly detect errors in 217 files.