### 2016-05-10 OSDI/concl.tex
config: concl

\section{Conclusion}

In this paper, we introduce \app, a highly modular framework 
that allows automatic verification of configuration files.
\app employs a translator to parse a sample configuration dataset
into a well-structured and typed intermediate representation,
and then uses a learner module to derive rules, 
thus building a language model.
For a given configuration file we want to verify,
\app uses the generated language model to check
whether any rule is violated.
We evaluate \app using a real-world dataset~\cite{configdataset}
which contains 261 incorrect MySQL configuration files.
Our experimental result shows \app is able to
correctly detect errors in 217 files in that dataset.

Our \app prototype still has many limitations. 
For example, we cannot handle configuration errors that can be 
triggered during system execution time, 
and we cannot detect misconfiguration across software components
maintained by different parties. 
Nevertheless, we believe \app represent a practical path
toward automatic language-based configuration verification.