### 2016-05-10 OSDI/concl.tex
aa

\section{Conclusion}

In this paper, we introduce \app, a highly modular framework 
that allows automatic verification of configuration files.
\app employs a translator to parse a sample configuration dataset
into a well-structured and typed intermediate representation,
and then uses a learner module to derive rules, 
building a language model this way.
\app then uses this generated language model 
to verify the correctness of
a given configuration file. The verification is done by
checking whether any of learned rules is violated.
We evaluate \app using a real-world dataset~\cite{configdataset}
which contains 261 incorrect MySQL configuration files.
Our experimental result shows that \app is able to
correctly detect errors in 217 files in that dataset.
\com{
Our \app prototype still has many limitations. 
For example, we cannot handle configuration errors that can be 
triggered during system execution time, 
and we cannot detect misconfiguration across software components
maintained by different parties. 
Nevertheless, w} We believe that \app represent a practical path
toward fully automated language-based configuration verification.