### 2016-08-30 Travis/secs/abstract.tex
restructure for ieee

\begin{abstract}
Rule based classification is an effective machine learning technique that yields low misclassification rates.
However, building a rule based system requires manual creation of large databases of logical constraints.
We present a method to generate rule based systems from temporally structured data.
As an demonstration of this algorithm, we plan to implement a learner that automatically generates constraints for the TravisCI testing framework.
The algorithm will utilize Github commit histories to generate logical constraints that allow us to detect potential build errors without actually building, saving valuable programmer and server time.
\end{abstract}