### 2016-09-04 Travis/secs/abstract.tex
better flow and motivation

\begin{abstract}
In previous work we used version space learning for configuration file verification.
We extend that work which only works on correct examples, to work on temporally structured correct and incorrect examples by building an SMT formula to find a classification model.
As an demonstration of this algorithm, we plan to implement it to automatically verify TravisCI configurations.
The algorithm will utilize Github histories to allow us to detect potential build errors without actually building, saving valuable programmer and server time.
\end{abstract}