### 2016-11-10 Ennan-Proposal/proposal.tex

\documentclass[11pt]{article}
\usepackage[margin=1in,includefoot]{geometry}

\usepackage{color}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
%\usepackage{bnf}
\usepackage{local}
\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{fancyvrb}
\usepackage[margin=1in]{geometry}

\usepackage{listings}
\usepackage{balance}

\lstset{
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  frame=none,
  language=haskell,
  %identifierstyle=\color{identifierColor},
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  %captionpos=b,                    % sets the caption-position to bottom
  %commentstyle=\color{mygreen},    % comment style
  %frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  otherkeywords={*,let, Server, Replication, FaultGraph, rankRCG, print, fialProb, goal, ...},           % if you want to add more keywords to the set
  numbers=none,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  %title=\lstname                  % show the filename of files included with \lstinputlisting; also try caption instead of title
  mathescape=true,
  tabsize=3,
  literate=*{->}{{\textcolor{blue}{$\to$}}}{1}
           {<-}{{\textcolor{blue}{$\leftarrow$}}}{1}
}

\begin{document}
\input{macros}

%\input{overview}

\newpage
\pagenumbering{arabic}
\setcounter{page}{1}

\input{motivation}
\input{prelim}
\input{arch}
\input{language}
\input{solutions}
\input{travis}
\input{related}
\newpage

\setcounter{page}{1}
\bibliographystyle{plain}
\bibliography{os}

\end{document}