### 2016-11-10 Ennan-Proposal/solutions.tex

\section{Learning More Accurate Specification}
\label{sec:learnin}

\ruzica{this is actually an OSDI summary and learning from incorrect files}