### 2016-11-12 Ennan-Proposal/impact.tex

\section{Broader Impact}

If successful, \app could have the following broader impacts:

\begin{itemize}

\item {\bf Opening the door of language-based configuration verification.}
  Today's misconfiguration checking approaches are still using {\em ad hoc}
  algorithms to detect potential misconfiguration root causes. On the
  contrary, \app could open the door of configuration verification,
  which rigorously verifies whether the target configuration files meet
  specifications. In other words, \app enables users 
  to verify configuration files like verifying programs. Such a capability
  cannot be offered by any of existing efforts.

\item {\bf Preventing configuration errors without adding any additional
  burden to users.} Because \app employs automatic machine learning 
  algorithms to extract rules, users or system administrators do not need
  to manually write their specifications, which significantly reduce the
  burden of users and system administrators. Such a way not only
  makes the configuration verification process become much easier and more 
  usable, but also decreases the potential issues of writing 
  specifications.

\item {\bf Complementing post-failure error diagnosis efforts.}
  There have been many diagnosis and troubleshooting efforts to localizing
  the root causes of configuration errors after system failures occur.
  We call these approaches as post-failure error diagnosis efforts.
  On the contrary, \app offers a pre-failure error checking effort,
  which could be looked as a preliminary checking before installing
  systems. Because \app is able to find out many sophisticated 
  configuration errors, it removes many labors of 
  the post-failure error diagnosis tools. In addition, the results
  output by \app could be used as a guidance for all the post-failure
  error diagnosis approaches, making these approaches more efficient.

\item {\bf Benefiting other communities, \eg, network misconfiguration.} 
  Besides software misconfiguration, network community has the same 
  misconfiguration problems. In particular, many recent service outage 
  reports exposed network-level configuration error will lead to
  very critical service outages. \app could be extended to reason about
  network-level configuration errors.

\end{itemize}