### 2016-11-14 Ennan-Proposal/agenda.tex
stuff

\section{Project Timeline}

The proposed project is supposed to last three years. We have identified three main 
goals of the project, each of them taking approximately a year:
\begin{description}
\item[Language Model] 
Our language model has to be expressive enough to be able to express and detect 
many non-trivial misconfiguration cases.
We will, therefore, need to define a more comprehensive language model (than the one presented).  The configuration files of some of software systems, 
\eg, squid, are not limited to the keyword-value parameters; thus,
we need to extend the current language model.
\item [Learning Algorithms]
We outlined basic learning algorithms. However,
many details are left unspecified and finding the right thresholds will be one of the main tasks.
We plan to investigate some existing learning algorithms for programs, \eg,
Raychev {\em et al.}~\cite{raychev15predicting, raychev16learning}. 
The configuration files have a specific structure and directly applying 
existing rules will not be possible. Still we will try to adapt some of these 
algorithms and ideas
to our problem.
\item [Real-Wolrd Evaluation]
Finally, we will implement and evaluate \app on real world misconfigurations.
In addition to evaluating \app on TravisCI files, during the development process 
we will empirically  evaluate efficiency, usability, and
accuracy of \app. We plan to use the existing publicly available dataset of 
misconfigurations~\cite{configdataset}. The set is large enough to use some parts as a  training data and some parts for evaluation.
\end{description}