### 2016-11-14 Ennan-Proposal/prior.tex

\section{Prior Results from NSF Support}

The PI received NSF CAREER Award for synthesis in live programming environment
(CCF-1553168 ). In addition, the PI received some student travel grants to support students' visit 
to conferences or summer schools. The most relevant NSF support for this proposal is the
NSF award ``SHF: Medium: Collaborative Research: FRP for Real'',
(CCF-1302327); \$866,000 (10/01/13-09/30/17), which the PI overtook  from 
Paul Hudak, due to his untimely death. Intellectual Merit: The
proposed work will contribute to a better understanding of declarative models
of time and interaction. Broader Impacts: The proposed work will further broaden
the applicability of FRP, improving implementation making FRP more suitable for
compute-intensive applications, such as interactive 3D graphics and real-time audio
processing. As the part of that grant we also developed and published ConfigC
(acknowledgments to the grant were given). Publications: M. Santolucito, 
E. Zhai, R. Piskac: ``Probabilistic Automated Language Learning for Configuration Files''
(July 2016), CAV 2016, Pages 80-87, DOI: 10.1007/978-3-319-41540-6\_5.