### 2017-01-18 CAV16/roadmap.tex

\section*{Outline}

\begin{itemize}

\item Introduction:

\item Motivating examples: Put two or three misconfiguration
  examples here, and show why they are incorrect.

\item \app Design: 
  \begin{itemize}
  \item Present an architecture of \app with a fig. Briefly describe 
    how it works (step by step).
  \item Detail learner part
  \item Merge
  \item Check 
  \item Limitations
  \end{itemize}

\item Implementation and Evaluation:
  First, briefly describe how we implemented \app.
  Then, show a few results, \eg, benchmark including learning time,
  checking time (with different size of configuration file).
  Case study?  I am not sure.

\end{itemize}