### 2017-01-21 CAV17/secs/intro.tex
remove old - add strucure

\section{Introduction}
\label{sec-intro}

Our main contributions are as follows:

\begin{enumerate}

% Encore also used association rule learning, but did not describe why it is useful for verification in particular,
%   or how it contrasts with other ML techniques that are less appropriate for verification (NN).
\item Describing the logical foundation of using association rule learning to build a probabilistic specification for configuration files. 
      We contrast this with the application of other machine learning techniques, such as neural networks, in verification.

% We can analyze the rules we have learned as a graph to sort the errors, not just by weight, but also importance
% need one more thing we can do with this graph
% This is only possible when we have learned association rules, not possible with NN
\item Analyzing the learned model to further refine the generated specification. 
      We highlight why these analysis techniques are uniquely possible in an association rule learning approach.

% The problem is intractable for off-the-shelf data mining solutions (see Encore).
% In addition to the optimizations Encore used, we add new ones that make our tool even better.  
\item An open-source implementation, \app, of this approach, which uses various domain specific optimizations to make the solution efficient. 
      We also evaluate \app on a set of 256 real world benchmarks, and identify XX new misconfigurations errors on YY of these files.
\end{enumerate}

Configuration errors are one of the most important root causes of today's software system failures~\cite{xu15systems, yin11anempirical}.
In their empirical study, Yin {\em et al.}~\cite{yin11anempirical} report that about 31\% of system failures were caused by misconfiguration problems and only 20\% were caused by bugs in program code. 
Misconfigurations, result in various system-wide problems, such as security vulnerabilities, application crashes, severe disruptions in software functionality, and incorrect program executions~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.

While many efforts have been proposed to check, troubleshoot, diagnose, and repair configuration errors~\cite{attariyan10automating,su07autobash, whitaker04configuration},
  those tools mainly post-factum and cannot give the safety guarantees available from powerful static analysis tools in program verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15}.
Existing techniques for program verification do not fit this application domain, as configuration files often lack specifications and structure.
There are rarely formal, or any, specifications on the constraints of the key values.
In the case of MySQL, the canonical specification is a 5240 page reference manual~\cite{mysql-manual15} to describe valid configurations and behavior.
We aim to automatically learn a formal specification the encodes this information by examining a set of example configuration files.

In previous work~\cite{santolucitoCAV}, we used a set of manually verified files by an domain expert order to learn specifications.
This approach is expensive and still prone to error by the expert.
Related work on configuration verification has used machine learning

While the grammar of configuration files tends to be simple (a series of key value pairs with some additional context), the keys and values themselves lack often structure and are untyped plain text.

   1) lack of a specification which would describe properties of configuration files, and 
   2) a program structure of configuration files -- they are mainly a sequence of entries assigning some value to system variables. 
The language in which configuration files are written does not adhere to a specific grammar or syntax. 
In particular, the entries in configuration files are untyped. 
Moreover, there are surprisingly few rules specifying constraints on entries and there is no explicit structure policy for the entries.
Thus, automated verification of configuration files would be highly desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

To overcome these obstacles, researchers proposed approaches based on
statistical analysis and learning%
~\cite{wang04automatic, zhang14encore, yuan11context}
that try to infer rules and policies about 
how configuration files are constructed.
Rather than explicitly specifying entries' types or rules, 
these efforts focused on learning policies from a sample dataset.
The learned rules are usually constructed in the following way: 
for every entry in a configuration file, 
they check if it deviates from a ``typical'' value, \ie, 
a value computed from a large training set 
consisting of configuration files.
If the entry is significantly different from a typical value, 
they suspect that it could be a potential configuration error.

The downside of these is that their learning approaches 
are limited to simplistic configuration errors, 
such as type errors and syntax errors, or they 
heavily rely on template-based inference~\cite{zhang14encore}. 
%As we mentioned earlier, 
%because configuration files are not written in a typed language 
%with a fixed syntax,
%type errors and syntax errors are relatively rare and easy to detect. 
Many sophisticated configuration errors, 
which are difficult to template in practice, 
cannot be detected in this way. 
As an illustration of such an error, consider entry ordering errors, 
which happens when some entries are inserted in a wrong order. 
For example, if in a PHP configuration file 
an entry {\tt extension = mysql.so} appears 
before {\tt extension = recode.so},  
it would lead to a crash error, where the Apache server cannot start 
due to the segmentation fault error. The correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical}.
These types of errors cannot be detected by existing
learning efforts, 
since it is hard to build a corresponding template~\cite{xu15systems}. 
%we would need to capture ordering relations of all the entries. 
%Some of them can appear in an arbitrary ordering, 
%and the resulting templates would be too large 
%and not efficient for a practical use.

%Offering automatic verification to configuration files -- like
%what we did to programs -- has been advocated as a reasonable means
%to check the correctness of configuration files of interest.
%Nevertheless, it is still an open problem, because 
%1) software configurations are typically written in poorly structured 
%and untyped languages, and 2) writing specifications or constraints
%for configuration verification is non-trivial in practice.

For practical purposes we use a real-world dataset~\cite{configdataset}. 
Every file in our dataset contains several errors, 
but they are typically different errors and only appear 
in a small percentage of files. Using that insight we were able, with the help of the probabilistic cutoff, to learn an
accurate set of rules. The rules, in general, specify which properties variables need to satisfy. One can see this learning process as a
way of deriving  
a specification for configuration files.  Once there is a specification, we can do formal verification. With these 
rules we can efficiently check the correctness of the configuration files
of interest and detect potential errors.

From a practical perspective, 
\app introduces no additional burden 
to the user: they can simply use \app to check for errors in their
configuration files. However, they can also easily extend the framework
themselves. The system is designed to be highly modular. If there is a
class of rules that \app is not currently learning, the user can develop
her own templates and learners for that class. The new learner can be
added to \app and this way it can check an additional new set of
errors.

\begin{enumerate}

\item We propose the first automatic configuration verification
framework, \app, that can learn a language model from a sample
dataset, and then use this language model to verify 
configuration files of interest.

\item \app proposes probabilistic types to assign a confidence 
distribution over a set of types to each entry, 
while generating the intermediate representation.

\item \app employs a collection of machine learning algorithms to 
enable powerful rule and constraint inference.

\item \app is capable of detecting various tricky errors that cannot
be detected by previous efforts,
including entry ordering errors, fine-grained value correlation errors, 
missing entry errors, and environment-related errors.

\item We implement a \app prototype and evaluate it by
conducting comprehensive experiments on real-world dataset.

\end{enumerate}