### 2017-01-23 CAV17/secs/concl.tex

\section{Conclusion}

In this paper, we introduce \app, a highly modular framework 
that allows automatic verification of configuration files.
\app employs a translator to parse a training set
of configuration files into an intermediate representation,
and then uses a learner module to derive rules.
\app finally uses post analysis module to refine these learned
rules, enabling users to verify the configuration files of
their interest.
Our evaluation based on a real-world misconfiguration
dataset~\cite{configdataset} shows that \app is able to
correctly detect configuration errors
including ordering, missing entry, integer correlation
and type errors.