### 2017-01-24 CAV17/secs/table2.tex
cav17: translator section

\begin{table*}[tbp]
\centering
\caption{Sampled misconfiguration files for error detection evaluation.}
\label{table-casestudy}
\begin{footnotesize}
\begin{tabular}{|c|c|c|c|}
\hline
{\bf Errors} & {\bf URLs} & {\bf Rank w/o} & {\bf Rank w/}\\ 
\hline
\hline
\multirow{3}{*}{MISSING ERROR: Expected "key-buffer"} & url & 7 & 8 \\
\multirow{3} & url & 9 & 10 \\ 
& url & 11 & 12 \\ 
\hline

\multirow{3}{*}{Ordering} & url & 7 & 8 \\ & url & 9 & 10 \\
& url & 11 & 12 \\  
\hline

\multirow{3}{*}{Type error} & url & 7 & 8 \\ & url & 9 & 10 \\  
& url & 11 & 12 \\
\hline

\multirow{3}{*}{fine-grained} & url & 7 & 8 \\ & url & 9 & 10 \\  
& url & 11 & 12 \\
\hline

\multirow{3}{*}{int correlation} & url & 7 & 8 \\ & url & 9 & 10 \\  
& url & 11 & 12 \\
\hline

\end{tabular}
\end{footnotesize}
\end{table*}