### 2017-01-24 CAV17/secs/table2.tex

\newcommand{\tablewidth}{6cm}
\definecolor{Gray}{gray}{0.80}
\newcolumntype{g}{>{\columncolor{Gray}}c}

\begin{table*}[tbp]
\centering
\caption{Sampled misconfiguration files for error detection evaluation.}
\label{table-casestudy}
\setlength{\tabcolsep}{0.5em}
\begin{footnotesize}
\begin{tabular}{|c|c|c|g|g|c|}
\hline
{\bf Errors} & {\bf URLs} & {\bf None} & {\bf PA} & {\bf PT} & {\bf PT$\land$ PA}\\ 
\hline
\hline
\multirow{3}{*}{\parbox{\tablewidth} {\scriptsize ORDERING ERROR: Expected ``innodb\_data\_home\_dir'' BEFORE ``innodb\_data\_file\_path''} }
& url & 12/12 & 3/12 & 5/5 & 3/5 \\
& url & 11/11 & 2/11 & 3/3 & 3/3 \\ 
& url & 9/9   & 3/9  & 4/4 & 3/4 \\ 
\hline

\multirow{3}{*}{\parbox{\tablewidth} {\scriptsize MISSING ERROR: Expected ``key\_buffer'' WITH [isamchk]} }
& url & 6/10 & 2/10 & 2/4 & 2/4 \\ 
& url & 2/3 & 3/3 & 2/3 & 3/3 \\
& url & 2/3 & 3/3 & 2/2 & 3/3 \\  
\hline

\multirow{3}{*}{\parbox{\tablewidth} {\scriptsize TYPE ERROR: Expected an integer for “slow\_query\_log”}}
& url & 32/34 & 1/34 & 5/7   & 1/7 \\ 
& url & 9/20  & 2/20 & 10/11 & 2/11 \\  
& url & 9/19  & 2/19 & 10/11 & 2/11 \\
\hline

\multirow{3}{*}{\parbox{\tablewidth} {\scriptsize FINE GRAINED ERROR: Expected \\ ``max\_connections'' * ``sort\_buffer\_size'' $\geq$ ``key\_buffer\_size''}}
& url & 30/34 & 18/24 & 6/7 & 3/7 \\ 
& url & 23/25 & 9/25  & 8/9 & 3/9 \\  
& url & 20/23 & 14/23 & 6/7 & 5/7 \\
\hline

\multirow{3}{*}{\parbox{\tablewidth} {\scriptsize INTEGER CORRELATION ERROR: Expected ``max\_allowed\_packet'' $<$ ``innodb\_buffer\_pool\_size''} }
& url & 29/32 & 8/32 & 11/14 & 4/14 \\ 
& url & 22/23 & 2/23 & 9/10  & 2/10 \\  
& url & 10/12 & 4/12 & 4/4   & 2/4 \\
\hline

\end{tabular}
\end{footnotesize}
\end{table*}