### 2017-01-25 CAV17/secs/concl.tex
cav17: stuff

\section{Conclusion}

In this paper, we introduce \app, a highly modular framework 
that allows automatic verification of configuration files.
\app employs a translator to parse a training set
of configuration files into an intermediate representation,
and then uses a learner module that employs
the association rule learning algorithm to derive rules.
\app finally uses rule graph analysis module to refine these learned
rules, enabling users to verify the configuration files of
their interest.
Our evaluation based on a real-world misconfiguration
dataset~\cite{configdataset} shows that \app is able to
correctly detect configuration errors
including ordering, missing entry, integer correlation
and type errors.