### 2017-01-25 CAV17/secs/abs.tex
autocommit

\begin{abstract}

System failures resulting from configuration errors are one of the major reasons for compromised reliability of today's software systems. 
Although many techniques have been proposed for configuration error detection, these approaches mainly can only be applied after an error has occurred. 
Verifying configuration files is, nevertheless, a challenging problem, because 
  1) software configurations are typically written in poorly structured and untyped ``languages'', and 
  2) specifying rules for configuration verification is challenging in practice.
This paper presents \app, a verification framework for general software configurations.
Our framework works as follows: in the pre-processing stage, we first automatically derive a specification.
Once we have a specification, we check if a given configuration file adheres to that specification.
The process of learning specification works through three steps.
First, \app parses a training set of configuration files (not necessarily all correct) into a well-structured and probabilistically-typed intermediate representation.
Second, based on the association rule learning algorithm \app learns rules from these intermediate representations. 
These rules establish relationships between the keywords appearing in the files. 
Finally, \app employs rule graph analysis to refine the resulting rules. 
\app is capable of detecting various configuration errors, including ordering errors, integer correlation errors, type errors, and missing entry errors. 
We evaluated \app by verifying public configuration files on Github, and we show that \app can detect known configuration errors in these files.

\end{abstract}