### 2017-04-15 PLDI/secs/concl.tex
fix typos

\section{Conclusion}

In this paper, we introduce \app, a highly modular framework 
that allows automatic verification of configuration files.
The main problem for verification of configuration files is their lack
of specification, so they are not a traditional target area
for formal methods.
Inspired by the association rule learning algorithm, \app learns 
a set of rules which describe properties and relations between keywords appearing in configuration files. These rules corresponding to the specification.
Our evaluation, based on real-world examples, shows that \app is able to
correctly detect and report configuration errors
including ordering, missing entry, integer correlation,
and type errors.