### 2017-05-06 ISSTA17/secs/intro.tex

\section{Introduction}

%what is the problem
Configuration files provide easy access to the most critical parameters of software systems in order to quickly tune the behavior of a system.
This expressive power in a single file creates a large surface for potential errors that can degrade performance or take down systems entirely.
The verification task for configuration files cannot be addressed with traditional software analysis techniques because they take a less structured form than a program.
As opposed to programs, configuration files rarely have formal specifications of a correctness, 
  the source code of the underlying system may not be available, 
  and the configuration may depend on the external environment such as the network or available hardware.
To overcome the unique challenges in this domain, we propose a method for automated configuration file verification using a data set of labeled examples of correct and incorrect configurations.

%What have we already done (labeled correct)
In our previous work~\cite{Santolucito2016} we built the tool ConfigC to verify configuration files from a set of correct examples.
ConfigC takes as a training set a set of correct configuration files and builds a set of rules describing a language model.
This model is then used to check if a new configuration file adheres to those rules.
If a file is incorrect, the tool identifies the location of error and reports what is incorrect.
For example, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so"} before \texttt{"extension recode.so"}.
While ConfigC's results are promising, the main weakness is that the learning can only use correct configuration files.
Additionally, there is a relatively high false positive rate (marking correct files as incorrect).

%What do we 'plan' to do to extend this (labeled correct and incorrect)
Here we propose to extend that work, which only works on correct examples, to work on both correct and incorrect examples by building an SMT formula to find a classification model.
By learning over both correct and incorrect examples, our hypothesis is that we will be able to decrease the high false positive rate observed in ConfigC, making the tool more useful to end users.
In order to extend ConfigC, we first provide a new description of ConfigC in terms of version space learning.
This description allows us to understand how the error reports are are produced in ConfigC.
Any extension to ConfigC must maintain the \textit{legibility} of learning - the tool must not only classify configuration files as correct or incorrect, but also identify the contributing factors (misconfigurations) to that classification.

%legibility issue
Legibility is a difficult property to achieve in machine learning and an active area of research~\cite{lei2016rationalizing}.
Although machine learning has been used in various software analysis techniques, such as programming-by-example \cite{lau2000version} and invariant synthesis \cite{garg2014ice}, this requires careful selection of the algorithms.
<any popular machine learning algorithms, such as neural nets and n-gram models, notoriously lack legibility and so are not well-suited to to provide simple justifications for their classification results.
The lack of justification for the results limits the applicability to software analysis, where providing feedback to a user is critical.
For example, in the case of error detection, the system should not only report which files have errors, but also locate the errors so a user may fix them.
By using non-probabilistic version space learning, we expect to have a highly level of legibility in our learning results.

%correctness issue
Additionally, probabilistic approaches to machine learning cannot be guaranteed to be complete, that is they are not guaranteed to always find an error if one exists.
We use version space learning to address the need for a formal completeness guarantee in an automated model generation for verification.
Version space learning is a technique for logical constraint classification~\cite{mitchell82}.
As a logical constraint system, version space learning not only provides justification of its classification results, but is has the foundation to build a completeness proof without the need for probabilistic thresholds.

Additionally, our extension takes advantage of a commonly available, but underutilized structure of training sets for software analysis.
While the previous training sets for ConfigC were an unsorted sampling of configuration files, typically we can obtain training sets that have a structure based on version history.
Because most code does not exist in isolation, but changes over time with development, any training set of code from a version control system, like Github, has this rich temporal structure. 
This structure is a partial order over time, and can be used by our proposed algorithm for more effective learning.

%evaluation plans
To test this approach in practice, we plan to implement our algorithm to check for TravisCI\footnote{\url{http://www.travis-ci.com}} configuration errors.
TravisCI is a continuous integration tool connected to Github that allows programmers to automatically run their test suite on every code update (commit).
A user adds a configuration file to the repository that enables TravisCI and specifies build conditions, such as which compiler to use, which dependencies are required, and a set of benchmarks to test.
This ensures the tool can always be automatically built correctly on a fresh machine.
The algorithm will utilize the temporal ordering on code in Github histories to allow us to detect potential build errors without actually building, saving valuable programmer and server time.

A recent usage study~\cite{API} of TravisCI found that 15-20\% of failed TravisCI builds are due to "errors" - which is the TravisCI name used to mean the configuration file was malformed and the software could not even be built.
Using the data from \cite{API}, we can also learn that since the start of 2014, approximately 88,000 hours of server time was used on TravisCI projects that resulted in an error status.
This number not only represents lost server time, but also lost developer time, as programmers must wait to verify that their work does not break the build.
If these malformed projects could be quickly statically checked on the client side, both TravisCI and its users could benefit.

Our main contributions are as follow:

\begin{enumerate}

\item We give a description of our tool, ConfigC, in the context of version space learning.
\item We propose a new algorithm to handle both incorrect and correct training data, as well as the internal structure of a training set.
\item We propose a real-world application of this approach, and outline the steps needed for an implementation.

\end{enumerate}