### 2017-06-26 ISSTA17/secs/vspace_verify.tex

\section{Version Space for Verification}

Version space learning builds a logical constraint model for binary classification, which we use to test a configuration file for membership in the set of all correct files~\cite{mitchell82}.
Traditional version space learning builds a model that tests membership using a series of disjunctions from a set of predefined hypotheses.
In ConfigC, we instead restrict the model to use a series of conjunctions.
In other words, rather than describing a correct file by allowable traits, we describe the required traits. 
This way, ConfigC can not only flag misconfigurations, but also give the points of failure for non-membership.
We now show in detail how this process works.

Our definition of a configuration file, $C$, is a file that can be transformed to an ordered list of $(keyword, value)$ pairs, called $Line$s.
ConfigC builds a model for a single file, $M(C)$, that is the set of all possible relations between the lines in the file.
A relation is described by user provided templates.
A template is a function taking some number of lines and determining a relation on the keywords of the lines ($Line^{n} \rightarrow Rel_{k1,..,kn}$).
As an example, ConfigC might derive the relation \texttt{extension mysql.so} comes before \texttt{extension recode.so} by using the ordering template.
ConfigC includes default templates for ordering, integer relations ($<,=,>$), type errors, and missing entries (two keywords must appear together irrespective of order).

It is assumed if a file is correct, all relations in that file are in the set of necessary relations, $Nec$, for any other file to be correct ($Correct(C) \implies \forall r \in M(C), r \in Nec$).
In this way, the initial model is built by creating the strongest conditions for a correct file, called the \textit{specific boundary} in version space learning.
This model is then iteratively relaxed as more examples are seen, a process called \textit{candidate elimination}.
To relax the model, two sets of relations from two files are merged into a single consistent set.

When the system is run on a new file, one of the relations from the templates will not be satisfied.
To provide an error message, the system only needs to report which relation is not satisfied.

Since we maintain the strongest condition for correctness, this approach will identify many correct files as incorrect, which we call a high false positive rate.
This is a problem when a user is then asked to manually review many files for errors, when the files are in fact correct.
However, the benefit to this approach is that, given the coverage of the training set, every misconfiguration will be detected.